\chapter{Beleidsinformatie Betuweroute \cite{beleidsinformatie-betuweroute}}
\section{Notes}
% Relevant notes
This is a study, executed by the Dutch audit office called \emph{De Rekenkamer}. 
The subject of this study is the process regarding the decision making of the \emph{Betuweroute}, or \emph{Betuwelijn} as it is called nowadays. 

This report is presented to the Dutch House of Representatives (\emph{tweede kamer}) while the project was still underway. 
The process when the project was started is described, and a number of remarks have been made.
The preparation of the government was lacking, there have been no really independent studies by experts. 
Basic information used during the decision making has also not been verified. 

A base decision has also been made during the start of the project. 
The incentive has been voiced to view the Betuwelijn as a preferable project\footnote{Preferable over road or water transportation.} 
However, alternatives were implicitly discarded after this preference had been voiced. 
This was partly because critical reports were ignored.

The document provides a critical, thorough review on the state of the subject as it was in the year 2000.

\section{Questions}
% Relevant questions
These questions are also answered within the document.
\begin{itemize}
  \item Why where the initial decisions about the Betuwelijn made?
  \item Based on what information have these decisions been made?
  \item Is the process of decision making correct in hindsight?
  \item How have decisions been modified in order to adhere environmental complaints?
  \item Have growth prognoses been set in a realistic manner?
\end{itemize}

\section{Review}
% Optional: review why fellow students have to read it
If you want to get a thorough and early insight in the process on the Betuwelijn, this is one of the two places to be. 
This document is as very clear and shows a lot of the mistakes that have been made in the early phases of the process. 
It is also very interesting to see some of the mistakes -- that have been made and pointed out in this document -- being repeated in later phases of the project.

It is a though read, but if interesting sections are read this is a really worthwile document.
